---
output:
  pdf_document: default
  html_document: default
editor_options: 
  markdown: 
    wrap: 72
---
# [Lab Session 4 (Week 12)]{style="color: rgb(61, 100, 200);"} {-}

## [What we have learnt so far]{style="color: rgb(61, 180, 140);"}

In the previous two weeks (Week 11 and Week 12), we have covered descriptive statistics, focusing on **measures of central tendency** and **measures of dispersion**. In this week's lab session we will bring everything together and we will also expand our work with skills that will allow us to present the above information visually.

## [Learning objectives]{style="color: rgb(61, 180, 140);"}  

* **Calculating measures of dispersion**
* **Installing the package [tidyverse]{style="color: dodgerblue;"}**
* **Install the package [ggpubr]{style="color: dodgerblue;"}**
* **Creating boxplots**
* **Creating barcharts that will also include statistical figures** 


## [Let's get started]{style="color: rgb(61, 180, 140);"}

**If you are working on RStudio Cloud skip to step 1 below**  

1. For this lab session you will need to create a new project in RStudio. Run RStudio and create a new project by clicking on File >> New Project ... .Then select New Directory and then New Project. Enter a Directory name. As mentioned in previous weeks, we recommend naming the folder by week, so "Week 12" would be ideal. Click **Browse** and make sure you place that folder in your "PSYC3000" folder that you created on your Desktop (if you have moved your work onto the z drive, then make sure to save your project there instead). Click "Create Project" to finalise creating your new project.  
2. Now create a new script from File >> New File >> R Script. Save your new script under the name **Week 12**.  
3. Our first step would be to install the package **[ggpubr]{style="color: dodgerblue;"}**. This package allows us to create customisable plots that we can include in our documents. This is useful for future assignments or publications. Type the following at the top of your script and run the line of code in order to install **[tidyverse]{style="color: dodgerblue;"}** and **[ggpubr]{style="color: dodgerblue;"}**.   


**install.packages("tidyverse")**  

**install.packages("ggpubr")**

One of the built-in datasets that R offers for free is called **[ToothGrowth]{style="color: FireBrick;"}**. 
The **[ToothGrowth]{style="color: FireBrick;"}** data set contains the result from an experiment studying the effect of vitamin C on tooth growth in 60 Guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, (orange juice (OJ) or ascorbic acid (a form of vitamin C and coded as VC).
Even though this dataset is not the result of Psychological research, it does have the structure of a data file from Psychological research where we administer an intervention in three different dosages and two different methods. We have worked with this dataset before but this time we will expand our work with more descriptive statistics and plots.  
 
 
## [Range, Interquartile Range, Median, and Relevant Plots]{style="color: rgb(61, 180, 140);"}
```{r message=FALSE, warning=FALSE, r,echo=FALSE}
# Set so that long lines in R will be wrapped:
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)
```
   

In your new script type and run the following:  
```{r echo=TRUE, message=TRUE, warning=FALSE}
library(tidyverse)
library(ggpubr)
data("ToothGrowth")
```

This informs R that we want to work with the package **[ggpubr]{style="color: dodgerblue;"}** and the built-in dataset called "**[ToothGrowth]{style="color:Firebrick;"}**". If you have a look at your Environment, under the section called **Data**, you will find an object called **[ToothGrowth]{style="color:Firebrick;"}**. You will also see some additional information, it has 60 observations and 3 variables. If you also move your mouse pointer on top of **[ToothGrowth]{style="color:Firebrick;"}** and leave it there you will get a tooltip window informing you that this is a **data.frame**. For now, you can consider a data.frame as a table that contains information for more than one variables. If you click once on **[ToothGrowth]{style="color:Firebrick;"}** then RStudio will open and display this data.frame for you. You can see it in a new tab that opened next to your script. You can see it has three columns, one per variable. The top row includes the variable names. **[len]{style="color:darkcyan;"}** is the variable that contains information on the tooth growth, **[supp]{style="color:darkcyan;"}** is the variable that contains information on the vitamin supplement used, and **[dose]{style="color:darkcyan;"}** is the variable that describes the dosage that was administered. 

If you type and run each of the following commands one at a time you will see each variable displayed in your console. This is pretty similar to what we did in **Week 3**.

```{r collapse=TRUE}
ToothGrowth$len

ToothGrowth$supp

ToothGrowth$dose
```

We will now start using our newly acquired knowledge of **measures of dispersion** and expand our skill-set. First, we can get all the basic descriptive information about our dataset with just one command line **[summary()]{style="color: dodgerblue;"}**

```{r }
summary(ToothGrowth)
```

You can see that the **[summary()]{style="color: dodgerblue;"}** functions returns the min, max, mean, median, Q1 and Q3 with just one execution. However, there are also some differences. Our variable **[len]{style="color:darkcyan;"}** is a numerical variable and that is why we get all this information. Where as our variable **[supp]{style="color:darkcyan;"}** is a factor with 2 levels and we only get a summary of how many cases we have in each level of the variable. In this case we see 30 for OJ and 30 for VC. As it was mentioned in the methods lecture of week 11, some variables are measurement variables whereas others can be manipulated variables. The variables we manipulate are known as **Independent Variables** (**IV**). When we manipulate variables then the measurement variables are our **Dependent Variables** or **DV**.

Manipulated variables have to have at least two levels. For example the **[supp]{style="color:darkcyan;"}** variable is a variable that we manipulated where we assigned **participants** randomly to one of the two conditions, **levels**. These two levels were VC and OJ. Similarly, **[dose]{style="color:darkcyan;"}** is also a variable we manipulated and it has three levels, **0.5**, **1.0**, and **2.0**.   

In your environment, you should be able to see an object called **[ToothGrowth]{style="color:firebrick;"}**. If you click on the little blue circle with the white triangle, you will get more information about **[ToothGrowth]{style="color:firebrick;"}**. We see that **[len]{style="color:darkcyan;"}** is numerical, which is correct as this was our **DV**, we also see that **[supp]{style="color:darkcyan;"}** is a factor with 2 levels. Finally. we see that **[dose]{style="color:darkcyan;"}** is also numerical, this is not correct. R does not always know/recognise the type of variables we are working with. In this case we have to inform R that **[dose]{style="color:darkcyan;"}** is also a factor. We will do that using the function **[as.factor()]{style="color:dodgerblue;"}**.

```{r }
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
```


With the above code we informed R that the variable **[dose]{style="color:darkcyan;"}** that is part of the **[ToothGrowth]{style="color:firebrick;"}** dataframe is a factor. Let's run the summary again and see whether the information we are getting is now different.

```{r }
summary(ToothGrowth)
```

You see now that we no longer get any numerical descriptive results. We only get a count of how many cases are in each level of the factor. This means that now R has properly identified our IVs. **When you work with data make sure your first step is to always check the variable and assign factors to your IVS**. Since **[len]{style="color:darkcyan;"}** is our only DV then we will focus on this one for our descriptives today. The following functions are self-explanatory, make sure to also add some comments in your code so that you can refer back to your script at later times.

```{r collapse = TRUE}

mean(ToothGrowth$len)
median(ToothGrowth$len)
range(ToothGrowth$len)
IQR(ToothGrowth$len)
var(ToothGrowth$len)
sd(ToothGrowth$len)

```


## [Exploring data using boxplots]{style="color:rgb(61, 180, 140);"}
One of the best plots we can create in order to visualise the *spread* of a variable is a **boxplot**. Boxplots contain information on the minimum and maximum values. They also include Q1, median, and Q3. Let us have a look.

```{r}
library(ggpubr) ### do not type this line if you already included at the top of your script
ggboxplot(ToothGrowth$len)
```

As you can see the boxplot is comprised of a **box** and two **lines**. The lower point of the line represents the min and the higher point represents the max. The horizontal line inside the box represents the median, and the lower and upper sides of the box represent the Q1 and Q3 respectively. Boxplots are even more informative if we use them to plot different variables, or create different boxplot per level of a factor. For example: 

```{r}
ggboxplot(ToothGrowth, x="dose", y = "len")
```

So now we can see one boxplot per level of **[dose]{style="color:darkcyan;"}**. Visually, the three doses seem to differ from each other. In the second term we learn how we can investigate whether these differences are meaningful or a product of chance. Back to our dataset now, we also have a second IV called **[supp]{style="color:darkcyan;"}**, we can ask **[ggboxplot]{style="color:dodgerblue;"}** to use that IV to create more boxplots. That way, we will have one boxplot per combination of the IV cells. We will explain this further below.

```{r}
ggboxplot(ToothGrowth, x="dose", y = "len", color = "supp")
```

So what we see here is six boxplots. With red colour we see all the box plots that correspond to the participants assigned to the **[OJ]{style="color:tomato;"}** level, with blue we see all the boxplots that correspond to the participants assigned to the **[VC]{style="color:cyan;"}** level. We still have the levels of the **[dose]{style="color:darkcyan;"}** on the x-axis. This representation helps us have some visual information about our data. For example, for the dose of 1mg we see that the VC boxplot is rather compressed, so we can say that our data in that combination had a rather small spread. Contrarily, the boxplot of the 2mg in the VC condition is much larger, so the data were spread much more in that condition. 

You can also swap position between **[dose]{style="color:darkcyan;"}** and **[supp]{style="color:darkcyan;"}** in the x variable and the color.

```{r}
ggboxplot(ToothGrowth, x="supp", y = "len", color = "dose")
```

Can you figure out what these boxplots represent? Instead of having three pairs, we now have two triads. Also, if you don't like white boxes you can use **fill** instead of **color**.

```{r}
ggboxplot(ToothGrowth, x="supp", y = "len", fill = "dose")
```



## [Exploring data using barcharts]{style="color: rgb(61, 180, 140);"}

While boxplots are great for visually exploring our data, they mainly represent the mean, the IQR, and the range. When we want to represent the mean of different groups we tend to use **barcharts**. For example: 

```{r}
ggbarplot(ToothGrowth$len)
```
   

This is a very basic version of a barchart that just includes all our DV measurements. Ideally we would want to see the mean of our DV for each level of an IV. For example the mean of **[len]{style="color: dark cyan;"}** per level of **[dose]{style="color: dark cyan;"}**.

```{r}
ggbarplot(ToothGrowth, x="dose", y = "len")
```
    

This is a very basic barchart and it also contains all individual measurements points as horizontal lines. We are interested in the mean.

```{r}
ggbarplot(ToothGrowth, x="dose", y = "len", add = "mean")
```
   

It would also be great to see the standard deviation of each group. We can do that by using **[add = "mean_sd"]{style="color: Violet;"}** instead of just **[add = "mean"]{style="color: Violet;"}**.  
```{r}
ggbarplot(ToothGrowth, x="dose", y = "len", add = "mean_sd")
```
   

These vertical lines are called **error bars** with the upper part being called **upper error bar** and the lower part **lower error bar**. In this case the upper and lower error bar is equal to the standard deviation. Furthermore, all the tricks we used for the boxplot can also be used here. For example:

```{r}
ggbarplot(ToothGrowth, x="dose", y = "len", fill = "supp", add = "mean_sd")
```
   

This does not really look great. This is because **[ggbarplot]{style="color: dodgerblue;"}** drew each bar on top of each other. We would rather have them next to each other. We need to add an [argument]{style="color: Violet;"} that will move the bars next to each other. Below, we use the **[position = position_dodge()]{style="color: Violet;"}** argument to do this:

```{r}
ggbarplot(ToothGrowth, x="dose", y = "len", fill = "supp", 
          position = position_dodge(0.7) , add = "mean_sd")
```

Both **[ggboxplot]**{style="color: dodgerblue;"} and **[ggbarplot]{style="color: dodgerblue;"}** come numerous more **arguments**. If you ever want to see a breakdown of a given function (including what arguments you can use), you can use the **[help]{style="color: dodgerblue;"}** function (see below) or google them to see what other options you have. We will also be using them a lot in the coming weeks so we will learn quite a few more arguments.

e.g.,

**[help(ggbarplot)]{style="color: dodgerblue;"}**

## **[Exercises]{style="color: rgb(180, 60, 140);"}**

The following graphs were made using **[ggboxplot]{style="color: dodgerblue;"}** and **[ggbarplot]{style="color: dodgerblue;"}**. We used the dataset **[mtcars]{style="color: firebrick;"}**. What you will have to do is use the right code to produce exactly the same graphs as shown in the following exercises. When you load the **[mtcars]{style="color: firebrick;"}** dataset, the variables **[cyl]{style="color: darkcyan;"}** and **[am]{style="color: darkcyan;"}** are numerical. **MAKE SURE YOU CONVERT THEM TO FACTOR AS YOUR FIRST STEP**. Look carefully in each graph so you can identify the variables that you will have to use to generate them.

### [Exercise 1]{style="color: rgb(180, 60, 140);"}

```{r echo=FALSE}
library(ggpubr)
data(mtcars)

mtcars$cyl <- as.factor(mtcars$cyl)
mtcars$am <- as.factor(mtcars$am)

ggboxplot(mtcars, x = "cyl", y = "disp")

```
   
### [Exercise 2]{style="color: rgb(180, 60, 140);"}
```{r echo=FALSE}
ggboxplot(mtcars, x = "cyl", y = "mpg", fill = "am")
```
   
### [Exercise 3]{style="color: rgb(180, 60, 140);"}

```{r echo=FALSE}
ggbarplot(mtcars, x = "am", y = "mpg", fill = "cyl", position = position_dodge(1), add = "mean_sd")
```
    
### [Exercise 4]{style="color: rgb(180, 60, 140);"}
```{r echo=FALSE}
ggbarplot(mtcars, x = "cyl", y = "mpg", color = "cyl", position = position_dodge(1), add = "mean")
```
    
### [Exercise 5]{style="color: rgb(180, 60, 140);"}
```{r echo=FALSE}
ggbarplot(mtcars, x = "cyl", y = "wt", fill = "am", position = position_dodge(0.5), add = "mean_sd")
```

.

.

.

.

.

.

.

.

.

.

.

## [Solutions: SPOILER ALERT]{style="color: rgb(61, 180, 140);"}
### [Exercise 1]{style="color: rgb(180, 60, 140);"}

```{r eval=FALSE}
library(ggpubr)
data(mtcars)

mtcars$cyl <- as.factor(mtcars$cyl)
mtcars$am <- as.factor(mtcars$am)

ggboxplot(mtcars, x = "cyl", y = "disp")

```
   
### [Exercise 2]{style="color: rgb(180, 60, 140);"}
```{r eval=FALSE}
ggboxplot(mtcars, x = "cyl", y = "mpg", fill = "am")
```
   
### [Exercise 3]{style="color: rgb(180, 60, 140);"}

```{r eval=FALSE}
ggbarplot(mtcars, x = "am", y = "mpg", fill = "cyl", position = position_dodge(1), add = "mean_sd")
```
    
### [Exercise 4]{style="color: rgb(180, 60, 140);"}
```{r eval=FALSE}
ggbarplot(mtcars, x = "cyl", y = "mpg", color = "cyl", position = position_dodge(1), add = "mean")
```
    
### [Exercise 5]{style="color: rgb(180, 60, 140);"}
```{r eval=FALSE}
ggbarplot(mtcars, x = "cyl", y = "wt", fill = "am", position = position_dodge(0.5), add = "mean_sd")
```
