---
output:
  pdf_document: default
  html_document: default
---
# Lab Session 7 (Week 16) {-}

## Learning Objectives
In this week's session we will be focusing on some "guess work". You will be given a partial script and some graphs. Your task will be to complete the code in order to produce the end-graphs on your own. These tasks will prepare you for the graphs you will need to create on your own for your report.
  

## To start with. 
First, start RStudio and create a new project. Your project should be named "Week 16 lab"

Go on moodle in week 9 in the lab session and download two files.  
1. "week9data.csv"   
2. "week9script.R"  

Make sure you save both files in the folder you just created for your new project.  
Once you have that then load the R script and execute **only the first 6 lines**.
In order to load the script go to File > Open File then select the "week9script.R" file and click on Open. If you can see the dataframe **df9** in your environment then you are all set to do the rest of the work in the lab.


## Lab task. The following tasks have to be completed in the lab   
### Part 1: Histograms

Go through your scripts now and try to complete the missing code in order to get the following graphs:

```{r echo=FALSE, message=FALSE, warning=FALSE}
## start of prep work
library(tidyverse)
library(ggpubr)

df9 <- read_csv("week9data.csv")
## end of preparation


## start of lab work
## compute the three averages of day, session, and task
## the following three lines of code are incomplete
## you will have to complete them on your own

df9 <- mutate(df9, day = (day1 + day2 + day3 + day4 + day5)/5)


df9 <- mutate(df9, session = (session1 + session2 + session3 + session4)/4)


df9 <- mutate(df9, task = (task1 + task2 + task3 + task4 + task5)/5 )



## in the next three lines amend the code to give the histograms
## exactly as they look in the given graph
## remember you maybe need to convert a variable to factor first

df9$group <- as.factor(df9$group)

gghistogram(df9, x = "day", bins = 8, 
            color = "group", fill = "group",
            add = "mean", 
            add.params = list(linetype = "solid", color = "group"),
            title = "Histograms per Group",
            xlab = "Day Performance",
            ylab = "Number of Participants")


gghistogram(df9, x = "session", bins = 5, 
            color = "group", fill = "group",
            add = "mean", 
            add.params = list(linetype = "solid", color = "group"),
            title = "Histograms per Group",
            xlab = "Session Performance",
            ylab = "Number of Participants")

gghistogram(df9, x = "task", bins = 12, 
            color = "group", fill = "group",
            add = "mean", 
            add.params = list(linetype = "solid", color = "group"),
            title = "Histograms per Group",
            xlab = "Task Performance",
            ylab = "Number of Participants")

```

### Part 2: Scatterplots

Now continue to the scatterplots

```{r message=FALSE, warning=FALSE, echo=FALSE}
p1 <- ggscatter(df9, x = "day", y = "session",
          title = "Scatterplot of Day by Session",
          xlab = "Day Performance",
          ylab = "Session Performance",
          cor.coef = TRUE) 
p1 + 
  font("title", size = 14, color = "red", face = "bold.italic") +
  font("xlab", size = 10, color = "orange") 
```

```{r message=FALSE, warning=FALSE, echo=FALSE}
p2 <- ggscatter(df9, x = "day", y = "task",
                title = "Scatterplot of Day by Task",
                xlab = "Day Performance",
                ylab = "Task Performance",
                cor.coef = TRUE) 
p2 + 
  font("title", size = 14, color = "blue", face = "bold") +
  font("xlab", size = 12, color = "green") +
  font("ylab", size = 12, color = "red") + 
  font("xy.text", size = 16, color = "gray", face = "bold")


```

```{r message=FALSE, warning=FALSE, echo=FALSE}
p3 <- ggscatter(df9, x = "task", y = "session",
                title = "Scatterplot of Task by Session",
                xlab = "Task Performance",
                ylab = "Session Performance",
                cor.coef = TRUE) 
p3 + 
  font("title", size = 14, color = "red", face = "bold") +
  font("xlab", size = 14, color = "blue") +
  font("ylab", size = 14, color = "blue") + 
  font("xy.text", size = 14, color = "red", face = "bold")


```

