---
output:
  html_document: default
  pdf_document: default
---
# SP300 Project A: Relationship between maths anxiety, test anxiety, and ...


## Your assignment (Deadline 18/12/2020 noon)

You will have to investigate for associations between **Maths Anxiety**, **Testing Anxiety**, and **two other constructs** of your choice from the list below. Which one is entirely up to you. However, make sure you link the four in your introduction so that your hypothesis and rationale behind the study are sound.  

You will have to produce a group report of 2,000 words (maximum). In this word count the references and the appendices are not included. In other words, the word count includes everything you have written from the abstract until the end of your discussion. This limit is the absolute maximum, which means you cannot go above it, you can however write below 2,000 words.  

On Moodle, there is a template file that you will have to use to write the report. You have to use that file as we want you to also fill in the first pages of the report. In these pages all group members will be listed and you will have to acknowledge all the members that actually attended and contributed. This is crucial as we want to be fair and only award students who have worked on the project.



## Psychological constructs in our study  

For the purposes of this report you will only need to work and analyse four variables. Two of them are compulsory, these are **Maths Anxiety** and  **Testing Anxiety**. From the following list you can choose any other two you want. You have to have a total of **four** variables in your work.  

1. **Maths Anxiety**  
2. **Testing Anxiety** 
3. Computer Programming Anxiety  
4. Studying Efforts  
5. Programming experience (experience)
6. Programming Progress (happy)
7. Test Progress  (tests)
8. Test Preparation  (test_hours)
9. Test Perseverance  (test_attempts)
10. Test Performance  (test_result)
11. Programming Skills (skills)





## Week 10 Session

You should have already chosen your four variables and today y ou should focus on completing all the analysis for them.
You should work together as a group and bring your dataset in a form where you can carry out correlational analysis. If you have createrd a project in RStudio in week 9 then you should continue working in that same project. This means that you have already downloaded the file **"project data.csv"**. **If not** then you should create a new project in RStudio and You should download the **"project data.csv"** file from moodle and save it in your project folder. You should then assign it to a dataframe and start working towards your analysis.Once the following is completed without any errors you should be good to start working within your group. If you have done that last week then you can skip this step.  


```{r echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(correlation)
library(ggpubr)


df <- read_csv("project data.csv")


```


 
## Data Preparation 

Before we are in a position to run our correlations we first have to calculate our questionnaires scores. In this study you completed a number of questions. We will not be analysing all of them. Instead we will only focus on. On Moodle, we have provided you with a pdf file that icludes most of the questions in this study. This file is called **"Qualtrics Survey.pdf"**. Make sure you download that file too and open it to familiarise yourself with the question. Next week we will be discussing the methods section, so this document will be useful for the completion of a number of sections in your report. **You will only need to prepare the variables you chose and not all of them**.

----  

## Maths Anxiety Scale (UK)  
This questionnaire is comprised of 23 items MAS_UK_1 to MAS_UK_23. It is scored on a 5-point Likert scale, *not at all (1)*, to *very much (5)*. You should calculate the sum for all the above items and name the variable **MAS**.  

----

## Test Anxiety Inventory  
This questionnaire is comprised of 20 items TAI_1 to TAI_20. It is scored on a 4-point Likert scale, *Almost Never (1)*, *Sometimes (2)*, *Often (3)*, and *Almost Always (4)*. You should calculate the sum of the above items and name the variable **TAI**.  

----

### Computer Programming Anxiety (CPAQ) (Brief Version)
This brief version of the questionnaire is comprised of 15 items. The full version of the questionnaire is much longer. However, for the purposes of our coursework this brief versions will be more than suitable. These 15 items are:   
1. GICS_1 to GIS_4 (Gaining Initial Computing Skills, GICS).   
2. SoC_1 to SoC_3 (Sense of Control, SOC).   
3. CSC_1 to CSC_4 (Computer Self Concept, CSC).   
4. SACS_1 to SACS_4 (State of Anxiety in Computer Situations, SACS).   


You should calculate four means. The **GICS**, **SOC**, **CSC**, and the **SACS**. Once you have these four means then you need to calculate a final variable called **CPAQ** that will be average of **GICS**, **SOC**, **CSC**, and the **SACS**.    

----

## Computer Programming Anxiety (CPAQ) Prior to joing our course
We also included a custom version of the above questionnaire where we asked you about your thoughts prior to joining our course. These items are similar with the above. We added the **_P** in the items names. So these items are GICS_P_1 to GIS_P_4, SoC_P_1 to SoC_P_3, CSC_P_1 to CSC_P_4, and SACS_P_1 to SACS_P_4. You should calculate four means. The **GICS_P**, **SOC_P**, **CSC_P**, and the **SACS_P**. Once you have these four means then you need to calculate a final variable called **CPAQ** that will be average of **GICS**, **SOC**, **CSC**, and the **SACS**. 


## Studying Efforts   
This one does not need any preparation and can be used as is. Refer to the questionnaire to see how the question was asked.   

## Programming experience (experience)  
This one does not need any preparation and can be used as is. Refer to the questionnaire to see how the question was asked.   

## Programming Progress (happy)  
This one does not need any preparation and can be used as is. Refer to the questionnaire to see how the question was asked.    

## Test Progress  (tests)  
This one does not need any preparation and can be used as is. Refer to the questionnaire to see how the question was asked. 

## Test Preparation  (test_hours)  
This one does not need any preparation and can be used as is. Refer to the questionnaire to see how the question was asked.  

## Test Perseverance  (test_attempts)  
This one does not need any preparation and can be used as is. Refer to the questionnaire to see how the question was asked.  

## Test Performance  (test_result)
This variable requires some preparation. Some students responded using 0-10 scale and some using 0-100. You should convert all answers to either 0-10 or 0-100 before you can use the variable.

## Programming Skills (skills)  
This one does not need any preparation and can be used as is. Refer to the questionnaire to see how the question was asked.  





```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

df <- mutate(df, GICS = (GICS_1 + GICS_2 + GICS_3 + GICS_4)/4)
df <- mutate(df, SOC = (SoC_1 + SoC_2 + SoC_3)/3)
df <- mutate(df, CSC = (CSC_1 + CSC_2 + CSC_3 + CSC_4)/4)
df <- mutate(df, SACS = (SACS_1 + SACS_2 + SACS_3 + SACS_4)/4)
df <- mutate(df, CPAQ = (GICS + SOC + CSC + SACS)/4)

df <- mutate(df, GICS_P = (GICS_P_1 + GICS_P_2 + GICS_P_3 + GICS_P_4)/4)
df <- mutate(df, SOC_P = (SoC_P_1 + SoC_P_2 + SoC_P_3)/3)
df <- mutate(df, CSC_P = (CSC_P_1 + CSC_P_2 + CSC_P_3 + CSC_P_4)/4)
df <- mutate(df, SACS_P = (SACS_P_1 + SACS_P_2 + SACS_P_3 + SACS_P_4)/4)
df <- mutate(df, CPAQ_P = (GICS_P + SOC_P + CSC_P + SACS_P)/4)

```


```{r eval=FALSE, include=FALSE}

df2 <- select(df, GICS, SOC, CSC, SACS, GICS_P, SOC_P, CSC_P, SACS_P)

correlation(df2)

summary(df2)
```
---- 



## Create a new dataframe that will only contain your 4 variables 
In order to make your work simpler you should create a new dataframe called `df2` that will only contain the four variables of your choice. You can do that using the function `select()`. We have used select() in the past so you refer to your notes on how to use itm, or use the help facility of RStudio.  


## Analysis needed for your report  


<<<<<<< HEAD
For your report you will need many different components that will go in different sections. Today we will focus on the **Results** section. For each of the **four chosen variables** you will need to create/report the following:  
=======
For your report you will need many different components that will go in different sections. Today we will focus on the **Results** section. For each of the *above calculated**four**  variables you will need to create/report the following:  


1. Mean, standard deviation, IQR and range  (this should go in the Results section)
2. A histogram for each variable including a line for the mean and a line for the median (This should go in the Appendices)

Furthermore, you will need to carry out the correlational analysis and include the following:  

<<<<<<< HEAD
1. All correlation coefficients between all your four variable as well as the *p-values*. hese should be reported in a table that will be formatted according to the American Psychological Association guidelines (APA 7). This should go in your results section
2.  We also want to see the scatterplots for all variable combinations (the four you chose). These should go in the Appendices. 

**NOTE: For all histograms and scatterplots we expect to see clear titles, x and y-axes labels, and legends. So make sure you format them correctly (according to APA) but also include the expected content.
=======
1. All correlation coefficients between all your four variable as well as the *p-values*. hese should be reported in a table that will be formatted according to the American Psychological Association guidelines (APA 7). This should go in your results section
2.  We also want to see the scatterplots for all variable combinations (the four you chose). These should go in the Appendices. 

**NOTE: For all histograms and scatterplots we expect to see clear titles, x and y-axes labes, and legends. So make sure you format them correctly (according to APA) but also include the expected content.




## Wrapping up  
Once you have finished all the above you should write a draft version of the results section so we can finalise them in the coming up Practical Session.





